---
title: "Case 7 Bazaar.com"
author: "Jonathan Ratschat"
date: "22 12 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#Libraries
library(readxl)
library(ggplot2)
library(reshape2)
library(dplyr)
```



```{r}
#Import data
Data <- read_excel("Bazaar data.xlsx")
head(Data)

#Transform variables
str(Data)
Data$Origin <- as.factor(Data$Origin)

#Create column with total clicks
Data$Total <- Data$Sponsored + Data$Organic

#Create column with indicator if brand keyword ad was used
Data$BrandKeyword <- ifelse(Data$Sponsored > 0, 1, 0)
Data$BrandKeyword <- as.factor(Data$BrandKeyword)

#Google Subset
dataGoogle <- subset(Data, Data$Origin == "Google")
```

# Exploratory analysis

```{r}
#Total clicks by Week and Origin
ggplot(Data, aes(x = Week, y = Total, fill = Origin)) +
         geom_bar(position = "dodge", stat = "identity") + 
         ggtitle("Total clicks by Week and Origin", subtitle = "No huge impact on total clicks when brand keyword ads are turned off") +
  labs(caption = "Ads were only used in the first nine weeks on Google (whole timespan on Bing).") +
xlab("Week") + 
  ylab("Total") + 
  theme(plot.title = element_text(color = "blue", face = "bold")) +
  scale_colour_manual(name="Origin") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12), labels = c("0", "2", "4", "6", "8", "10", "12")) +
  expand_limits(x = 0, y = 0)

```

```{r}
#Google clicks by week
dataGoogleLong <- melt(dataGoogle %>% 
                         select(-BrandKeyword, -Total), 
                       id = c("Week", "Origin"))

ggplot(dataGoogleLong, aes(x = Week, y = value, fill = variable)) +
         geom_bar(position = "stack", stat = "identity") + 
         ggtitle("Google clicks by Week", subtitle = "When sponsored ads are turned off, many clicks are substituted by organic clicks") +
xlab("Week") + 
  ylab("Number of clicks") + 
  theme(plot.title = element_text(color = "blue", face = "bold")) +
  guides(fill=guide_legend(title="Type")) +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12), labels = c("0", "2", "4", "6", "8", "10", "12")) +
  expand_limits(x = 0, y = 0)
```

# Regression analysis

Perform regression analysis to find out if BrandKeyword has a significant effect on log(Total). The log is chosen to make the results of Bing and Google more comparable.

```{r}
#Regression on Google Subset using BrandKeyword as independent variable
mod1 <- lm(log(Total) ~ BrandKeyword + Week, data = dataGoogle)
summary(mod1)
```

BranKeyword has only a small effect on log(Total) at a low significance 10% level.

# Difference-in-difference analysis

```{r}
#Regression on Data using interaction effect between week and BrandKeyword as independent variable
mod2 <- lm(log(Total) ~ Week*BrandKeyword + Origin, data = Data)
summary(mod2)
```

BrandKeyword and Week:BrandKeyword1 coefficients are not significant. Zero hypothesis that coefficients have no effect on dependent variable cannot be rejected.

# Conclusion

All regression results confirm the very small and statistically insignificant effect of paid search ads.