---
title: "Case 7 Bazaar.com"
author: "Jonathan Ratschat / Franziska Bülck"
date: "22 12 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Libraries

```{r}
#Libraries
library(readxl)
library(ggplot2)
library(reshape2)
library(dplyr)
```

# Prepare data

```{r}
#Import data
Data <- read_excel("Bazaar data.xlsx")
head(Data)

#Transform variables
str(Data)
Data$Origin <- as.factor(Data$Origin)

#Create column with total clicks
Data$Total <- Data$Sponsored + Data$Organic

#Create time indicator (before treatment = 0, after treatment = 1)
Data$Time <- ifelse(Data$Week < 10, 0, 1)
Data$Time <- as.factor(Data$Time)

#Data set used for analysis
head(Data)
str(Data)
```

# Exploratory analysis

```{r}
#Total clicks by Week and Origin
ggplot(Data, aes(x = Week, y = Total, fill = Origin)) +
         geom_bar(position = "dodge", stat = "identity") + 
         ggtitle("Total clicks by Week and Origin", subtitle = "Number of total clicks shows a decreasing trend") +
  labs(caption = "Ads were only used in the first nine weeks on Google (whole timespan on Bing).") +
xlab("Week") + 
  ylab("Total") + 
  theme(plot.title = element_text(color = "blue", face = "bold")) +
  scale_colour_manual(name="Origin") +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))

```

```{r}
#Google Subset
dataGoogle <- subset(Data, Data$Origin == "Google")

#Google clicks by week
dataGoogleLong <- melt(dataGoogle %>% 
                         select(-Total, -Time), 
                       id = c("Week", "Origin"))

ggplot(dataGoogleLong, aes(x = Week, y = value, fill = variable)) +
         geom_bar(position = "stack", stat = "identity") + 
         ggtitle("Google clicks by Week", subtitle = "When sponsored ads are turned off, many clicks are substituted by organic clicks") +
xlab("Week") + 
  ylab("Number of clicks") + 
   labs(caption = "Ads were only used in the first nine weeks on Google.") +
  theme(plot.title = element_text(color = "blue", face = "bold")) +
  guides(fill=guide_legend(title="Type")) +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```

# Difference-in-difference analysis

We perform difference-in-difference analysis to find out if using paid search ads (brandkeywords) has a significant effect on log(Total). The log is chosen to make the results of Bing and Google more comparable.

Two independent variables:
Time: before and after treatment (treatment: turn brandkeywords off)
Origin: treatment group and control group (treatment = Google, control = Bing)

```{r}
mod1 <- lm(log(Total) ~ Time*Origin, data = Data)
summary(mod1)
```

Time1:OriginGoogle coefficient is not significant. Zero hypothesis that the usage of brandkeywords has no effect on dependent variable cannot be rejected.

# Conclusion

The conducted difference-in-difference analysis confirms the statistically insignificant effect of paid search ads. Therefore, we recommend to stop spending money for brand keyword ads.
