---
title: 'Linear Regression: Direct Marketing'
author: "Jonathan Ratschat"
date: "1 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Regression: Direct Marketing

## Libraries

```{r}
library(readr)
```

## Load data

```{r}
Data <- read_csv("direct_marketing.csv")
```

## Look at data

```{r}
summary(Data)
str(Data)
```

## Transform variables

```{r}
Data$Age <- as.factor(Data$Age)
Data$Gender <- as.factor(Data$Gender)
Data$OwnHome <- as.factor(Data$OwnHome)
Data$Married <- as.factor(Data$Married)
Data$Location <- as.factor(Data$Location)
Data$History <- as.factor(Data$History)
Data$Children <- as.factor(Data$Children)
```

## What are the characteristics of the most attractive customer?

```{r}
mod1 <- lm(AmountSpent ~ Age + Gender + OwnHome + Married + Location + Salary + Children + Catalogs, data = Data)
summary(mod1)

mod2 <- lm(AmountSpent ~ Age + Gender + OwnHome + Married + Location + Salary + Children + Catalogs + History, data = Data)
summary(mod2)
```

Young age, LocationFar, High Salary, No children, No. of catalogs sent

## What is the expected amount of dollars spent of a customer with the following characteristics: young, female, home owner, single, far away living, $30.500 salary, 12 catalogs sent

```{r}
-6.586e+02 + 9.123e+01 + 5.061e+02 + 30500 * 2.213e-02 + 12 * 4.334e+01
```

## Would you rather target married or single customers?

It does not matter, since this variable is not significant. Therefore, zero-hypothesis that this variable has no effect on dollar amount cannot be rejected.

## Which customers spent more money than you expected?

```{r}
Data$Predicted <- predict(mod1, data = Data)

Data$difference <- Data$AmountSpent - Data$Predicted
```

Customers with values with positive difference spend more money than expected, while customer with negative values spend less money than expected.