---
title: "02 - Alumni Giving"
author: "Nils Gandlau"
date: "23 10 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)

dataOrig <- read_csv("C:/Users/nilsg/Google Drive/Studium/WS1920/analytics_customer_value_management/case_studies/02_alumni_giving/data_alumni_giving.csv")
data <- as.data.table(dataOrig)
```


## Data Preparation

```{r}
str(data)
```

### Data Types
Percentages are saved as characters. Reformat those to make them numeric.

```{r}
data <- data %>% 
  mutate(
    SFR = SFR,
    LT20 = as.numeric(str_extract(LT20, "[0-9]*")) / 100,
    GT50 = as.numeric(str_extract(GT50, "[0-9]*")) / 100,
    GRAD = as.numeric(str_extract(GRAD, "[0-9]*")) / 100,
    FRR = as.numeric(str_extract(FRR, "[0-9]*")) / 100,
    GIVE = as.numeric(str_extract(GIVE, "[0-9]*")) / 100,
  ) %>% 
  setDT()

str(data)
```

### Missing Values

There are no missing values.

```{r}
nas <- rbindlist(lapply(names(data), function(colname) {
  nMissingValues <- sum(is.na(data[, get(colname)]))
  return(data.table(feature = colname, number_of_nas = nMissingValues))
}))

nas
```


## Data Exploration

```{r fig.height=12, fig.width=10}
par(mfrow = c(4, 2))
plot(GIVE ~ ., data %>% select(-ID, -School))
```

One may identify the following trends:

* The larger the student-to-faculty ratio, the lower the giving rate
* The larger the share of classes with fewer than 20 students, the larger the giving rate
* The larger the share of classes with more than 20 students, the slightly lower the giving rate
* The larger the average six-year graduation rate, the larger the giving rate
* The more freshmen stay, the larger the giving rate

### Correlations of (numeric) features

```{r}
numericFeatures <- c("GIVE", "SFR", "LT20", "GT50", "GRAD", "FRR")
dataNumericFeatures <- data[, c(numericFeatures), with = F]

corMatrix <- round(cor(dataNumericFeatures), 2)
corMatrix
```

* `GRAD` and `FRR` have high (positive) correlation $0.93$. We may want to avoid using both in our models.
* `SFR` and `LT20` have medium-high (negative) correlation $-0.69$. We also note that the two variables have, when increased, an *opposing effect* on the target variable. Hence, when both were used in a regression model, their regression coefficients may be biased/untrustworthy due to multicolinearity.

## Regression

We deploy a standard linear regression, but exclude `FRR` since it is highly correlated with `GRAD`.

```{r}
linReg1 <- lm(GIVE ~ SFR + LT20 + GT50 + GRAD, data = data)
summary(linReg1)
```

We check whether including `FRR` over `GRAD` is better:

```{r}
linReg2 <- lm(GIVE ~ SFR + LT20 + GT50 + FRR, data = data)
summary(linReg2)
```

* We note that in the second regression, the $R^2$ is slightly lower, hence we will use the first one, where we included `GRAD` and not `FRR`

### Question 1

If School A's graduation rate `GRAD` is 10 (percentage) points higher (note that 10 percentage points equals 0.1 in our case, since we converted the percentages to be numbers in $[0, 1]$) than School B's, then School A can expect to have an *alumni giving rate* that is $0.257573 * 0.1 * 100 \approx 2.6$ percentage points higher than School B's.

The calculation is based on the following idea:

$$
\begin{aligned}
\Delta GIVE & =(w_0 +w_1 GRAD_2 + \dots) - (w_0 + w_1 GRAD_1 + \dots) \\
& = w_1 \cdot (GRAD_2 - GRAD_1) \\
& = w_1 \cdot 0.1
\end{aligned}
$$

### Question 2

The answer doesn't change compared to Question 1. 

Since we have assumed for each predictor a linear relationship with the target variable, the benefit of a marginal increase in `GRAD` is not dependent on the (absolute) level of any of the predictors. 

That is, regardless of the value of student-to-faculty ratio, our model says that increasing the graduation rate by 10 percentage points (or 0.1 in our case), always increases the expected giving rate by 2.6 percentage points.

### Question 3

```{r}
maxGIVE <- max(data$GIVE)
minGIVE <- min(data$GIVE)

rbindlist(list(
  data[GIVE == maxGIVE, c("School", "GIVE")],
  data[GIVE == minGIVE, c("School", "GIVE")]
))
```

* University of Notre Dame has the most impressive giving rate (41%)
* Three universities share the last place: San Diego State University, San Jose State University, University of South Alabama. Each with a giving rate of 2%.

### Question 4

School's feature vector:

* `GRAD = 0.67`
* `SFR = 17`
* `LT20 = 0.34`
* `GT50 = 0.23`
* `FRR = 0.77`

Note that in our regressions before, we haven't included both `GRAD` and `FRR` at the same time, since they were highly correlated. Hence, we have to run a new regression:

```{r}
linReg <- lm(GIVE ~ SFR + LT20 + GT50 + GRAD + FRR, data = data)
summary(linReg)
```


Given the school's feature vector, we can compute the expected giving rate that is predicted by our model:

$$
\begin{aligned}
GRAD & = w_0 + w_1 \cdot SFR+w_2 \cdot LT20 + w_3\cdot GT50 + w_4 \cdot GRAD + w_5 \cdot FRR \\
& \approx -0.1951-0.0011\cdot 17 + 0.1509\cdot 0.34 -0.0315 \cdot 0.23 + 0.1297 \cdot 0.67 + 0.2570 \cdot 0.77 \\
& = 0.11505
\end{aligned}
$$
So our model predicts a graduation rate of 11.505 %, which is greater than 8%.
