---
title: "CaseStudy-04: Retail Relay"
author: "Nils Gandlau"
date: "19 11 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tidyverse)
library(readxl)
library(magrittr) # for %<>% operator 

# Load data
dataMainFull <- as.data.table(read_xlsx(path = "retailRelay.xlsx", sheet = 1))
dataMainPilot <- as.data.table(read_xlsx(path = "retailRelay.xlsx", sheet = 2))
dataTransitionFull <- as.data.table(read_xlsx(path = "retailRelay.xlsx", sheet = 3))
dataTransitionPilot <- as.data.table(read_xlsx(path = "retailRelay.xlsx", sheet = 4))

# Rename Variables
dataMainPilot %<>% rename(OrderId = `Order Id`)

dataTransitionFull[, ':='(
  transProb = as.numeric(`Transition Probability`),
  purchaseOccasion = `Purchase Occasion`,
  avgBasketSize = `Average Basket Size`
)]
dataTransitionPilot[, ':='(
  transProb = as.numeric(`Transition Probability`),
  purchaseOccasion = `Purchase Occasion`,
  avgBasketSize = `Average Basket Size`
)]

dataTransitionFull <- dataTransitionFull[, c("purchaseOccasion", "transProb", "avgBasketSize")]
dataTransitionPilot <- dataTransitionPilot[, c("purchaseOccasion", "transProb", "avgBasketSize")]

# Transforming dates
dataMainFull[, ':='(orderYear = lubridate::year(OrderDate),
                    orderMonth = lubridate::month(OrderDate),
                    orderDay = lubridate::day(OrderDate))]
dataMainPilot[, ':='(orderYear = lubridate::year(OrderDate),
                    orderMonth = lubridate::month(OrderDate),
                    orderDay = lubridate::day(OrderDate))]
```

## Research Question

Are you happy with the development of the customer base?

## Data


## Transition

```{r}
dataTransitionFull[, label := "full"]
dataTransitionPilot[, label := "pilot"]
dataTransition <- rbind(dataTransitionFull, dataTransitionPilot)

# Compute average Transition
dataTransitionFull[, transProbPilot := dataTransitionPilot$transProb]
dataTransitionFull[, avgBasketSizePilot := dataTransitionPilot$avgBasketSize]
dataTransitionFull[, avgTransition := 0.5 * (transProb + transProbPilot)]
dataTransitionFull[, avgAvgBasketSize := 0.5 * (avgBasketSize + avgBasketSizePilot)]
dataTransitionAvg <- dataTransitionFull[, c("purchaseOccasion", "avgTransition", "avgAvgBasketSize")]
setnames(dataTransitionAvg, c("purchaseOccasion", "transProb", "avgBasketSize"))
dataTransitionAvg[, label := "average"]

dataTransition <- rbind(dataTransition, dataTransitionAvg)

head(dataTransition)
```

```{r warning=FALSE}
ggplot(dataTransition, aes(x = purchaseOccasion, y = transProb, color = label)) +
  geom_line(size=1) +
  geom_point() +
  theme_minimal() +
  xlim(c(1, 30)) +
  ylim(c(0, 1))
```

## Customer base

```{r}
# Can the same OrderId occur multiple times? => NO.
length(unique(dataMainFull$OrderId))
```


```{r}
# Construct purchaseOccasion
dataMainFull[, purchaseOccasion := frank(OrderId), by = UserId]
dataMainPilot[, purchaseOccasion := frank(OrderId), by = UserId]
```


```{r}
# How many users in pilot vs full study?
length(unique(dataMainPilot$UserId))
length(unique(dataMainFull$UserId))
```


```{r}
# Does dataMainFull include the same data for a customer?
y <- sapply(unique(dataMainPilot$UserId), function(userId){
  dt1 <- dataMainPilot[UserId == userId]
  dt1 <- as.data.frame(dt1[, c("UserId", "OrderId", "OrderDate")])
  
  dt2 <- dataMainFull[UserId == userId]
  dt2 <- as.data.frame(dt2[, c("UserId", "OrderId", "OrderDate")])
  
  return(nrow(dt1) == nrow(dt2))
})

all(y)
```


## Defining cohorts

What's the time span of full vs pilot

```{r}
min(dataMainPilot$OrderDate)
max(dataMainPilot$OrderDate)
cat("")
min(dataMainFull$OrderDate)
max(dataMainFull$OrderDate)
```

How many people are in January 2009, Februrary 2009, .... (in each month)

```{r}
dataMainFull %>%
  group_by(orderYear, orderMonth) %>% 
  distinct(UserId) %>% 
  count()
```

```{r}
dataMainPilot %>%
  group_by(orderYear, orderMonth) %>% 
  distinct(UserId) %>% 
  count()
```

### Cohorts

Cohort := customers that joined in the same month

We have 12+3=15 months in the data. So we will build 15 cohorts.

```{r}
# Assigning customers to cohorts 1,...,15
dataMainFull[, cohort := rep(0, nrow(dataMainFull))]
dataMainFull[, cohort := ifelse(orderYear == 2009, min(orderMonth), 0), by = UserId]
dataMainFull[, cohort := max(cohort), by = UserId]
dataMainFull[orderYear == 2010, cohort := ifelse(cohort == 0, min(orderMonth) + 12, cohort), by = UserId]
```

```{r}
# Number of customers by cohort
dataByCohort <- dataMainFull %>% 
  mutate(cohort = as.factor(cohort)) %>% 
  group_by(cohort) %>% 
  summarise(nCustomers = n_distinct(UserId))

ggplot(dataByCohort, aes(x = cohort, y = nCustomers, fill = cohort)) +
  geom_col() +
  geom_text(aes(label = nCustomers), vjust = -0.2)
```

```{r fig.width=12}
# Number of Customers by cohort and purchase Occasion (relevant for Aussagekraft of plot below)
dataByCohortPurchase <- dataMainFull %>% 
  mutate(cohort = as.factor(cohort)) %>% 
  group_by(cohort, purchaseOccasion) %>% 
  summarise(nCustomers = n_distinct(UserId)) %>% 
  filter(purchaseOccasion %in% c(0:40))

ggplot(dataByCohortPurchase, aes(x = purchaseOccasion, y = cohort, fill = nCustomers)) +
  geom_tile(color = "black") +
  geom_text(aes(label = nCustomers), size = 3.5) +
  scale_fill_gradient(low = "white", high = "darkgreen")
```


Average basket size by cohort 

```{r}
dataPlt1 <- dataMainFull %>% 
  group_by(cohort, purchaseOccasion) %>% 
  summarise(avgOrder = round(mean(TotalCharges, na.rm=T),0)) %>%
  ungroup() %>% 
  mutate(cohort = as.factor(cohort))

dataPlt1
```


```{r fig.width = 12}
# Note that we can't trust these averages because they're computed based on a single customer.... that's kind of stupid.!

# Average order-volume ($) by purchase occasion and cohort
ggplot(dataPlt1 %>% filter(purchaseOccasion %in% 1:40), aes(x = purchaseOccasion, y = cohort, fill = avgOrder)) +
  geom_tile(color = "black") +
  geom_text(aes(label = avgOrder), size = 3.5) +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  theme_minimal()
```


```{r}
ggplot(dataPlt1, aes(x = purchaseOccasion, y = avgOrder, fill = cohort)) +
  geom_area(color = "black", alpha = 0.8, position = "dodge")
```

```{r}
ggplot(dataPlt1, aes(x = purchaseOccasion, y = avgOrder, fill = cohort)) +
  geom_col(position = "stack", alpha = 1)
```

```{r}
dataPlt1 %>% filter(cohort %in% 14:15) %>% filter(purchaseOccasion %in% 30:60)
```

```{r}
max(dataPlt1$purchaseOccasion) 
```

```{r}
ggplot(dataPlt1, aes(x = purchaseOccasion, y = avgOrder)) +
  geom_line() +
  facet_wrap(~cohort)
```


#### Purchase quality by cohort

```{r}
dataByCohort <- dataMainFull %>% 
  mutate(cohort = as.factor(cohort)) %>% 
  group_by(cohort) %>% 
  summarise(
    nCustomers = n_distinct(UserId),
    maxPurchaseOccasion = max(purchaseOccasion),
    nPurchases = n_distinct(OrderId),
    cumulativeOrderSize = sum(TotalCharges),
    avgOrderSize = mean(TotalCharges),
    nPurchasesPerCustomer = round(nPurchases / nCustomers, 0),
    orderSizeDensity = round(cumulativeOrderSize / nPurchases / maxPurchaseOccasion / nCustomers, 4)
  )

dataByCohort
```

```{r fig.width=12}
# note: possibly add "percentile" => how many % of customers make 90% of orders? 
# include timespan => cohort 1 has more purchases because it's over a longer timespan

subtitle = "U = Number of Users
P = Number of Purchases 
PpC = P/U = Number of purchases per User"

# Average order volume by cohort
ggplot(dataByCohort, aes(x = cohort, y = avgOrderSize, fill = cohort)) + 
  geom_col() +
  geom_text(aes(label = paste0(nCustomers, " U \n", nPurchases, " P \n", nPurchasesPerCustomer, " PpC")), size = 3, vjust = -0.5) +
  geom_text(aes(label = paste0(round(avgOrderSize, 0), "€")), vjust = 1.5) +
  ylim(c(0, 90)) +
  theme_minimal() +
  ggtitle("Average Order-Size ($) for each cohort", 
          subtitle = subtitle)
```

For example, in cohort 15 there was a total of 2213€ made over 55 purchases, distributed over 9 purchase occasions and 37 customers. Hence the "density" is 2213/55/9/37=0.88. In other words, customers in cohort 15 seem to spend more in less occasions. 

```{r}
# € Density
ggplot(dataByCohort, aes(x = cohort, y = orderSizeDensity, fill = cohort)) +
  geom_col() +
  ggtitle("€ per Customer/Occasions/Purchases [Density] ")
```



## Higher aggregate cohorts (5-Month-Cohorts)

```{r}
dataMainFull[, cohort5month := ifelse(cohort %in% c(1:5), 1, ifelse(cohort %in% c(6:10), 2, 3))]

dataCohort5month <- dataMainFull %>% 
  mutate(cohort5month = as.factor(cohort5month)) %>% 
  group_by(cohort5month, purchaseOccasion) %>% 
  summarise(
    nCustomers = n_distinct(UserId),
    cumulativeOrderSize = sum(TotalCharges),
    avgOrderSize = mean(TotalCharges),
    stddevOrderSize = sd(TotalCharges)
    )
```


```{r fig.width=12}
# Average order-volume ($) by purchase occasion and cohort
ggplot(dataCohort5month %>% filter(purchaseOccasion %in% 1:15), aes(x = purchaseOccasion, y = cohort5month, fill = avgOrderSize)) +
  geom_tile(color = "black") +
  geom_text(aes(label = paste0(round(avgOrderSize,0), "€ \n", "sd=", round(stddevOrderSize,0), "\n","(N=", nCustomers, ")")), size = 3) +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  theme_minimal() + 
  ggtitle("Purchases of 5-month cohorts",
          subtitle = "sd = standard deviation of the order size (€) in the given tile
N = number of customers in the given tile")
```

## Quarterly Cohorts

```{r}
# Defining quarterly cohorts
dataMainFull[, cohortQuarter := ifelse(
  cohort %in% c(1:3),
  1, 
  ifelse(
    cohort %in% c(4:6),
    2,
    ifelse(
      cohort %in% c(7:9),
      3,
      ifelse(
        cohort %in% c(10:12),
        4,
        5))))]

dataCohortQuarter <- dataMainFull %>% 
  mutate(cohortQuarter = as.factor(cohortQuarter)) %>% 
  group_by(cohortQuarter, purchaseOccasion) %>% 
  summarise(
    nCustomers = n_distinct(UserId),
    cumulativeOrderSize = sum(TotalCharges),
    avgOrderSize = mean(TotalCharges),
    stddevOrderSize = sd(TotalCharges)
    )
```

```{r fig.width=12}
# Average order-volume ($) by purchase occasion and cohort
ggplot(dataCohortQuarter %>% filter(purchaseOccasion %in% 1:20), aes(x = purchaseOccasion, y = cohortQuarter, fill = avgOrderSize)) +
  geom_tile(color = "black") +
  geom_text(aes(label = paste0(round(avgOrderSize,0), "€ \n", "sd=", round(stddevOrderSize,0), "\n","(N=", nCustomers, ")")), size = 3) +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  theme_minimal() + 
  ggtitle("Purchases of quarterly cohorts",
          subtitle = "sd = standard deviation of the order size (€) in the given tile
N = number of customers in the given tile")
```

## Cohort size

Reconstructing cohort size by assuming sampled uniformly from all customers. Proportion in sample represents somewhat approx. the growth of number of customers.